---
title: "Murder Gun"
author: "Ladan Foroughi"
date: "10/05/2021"
output:
  pdf_document:
    latex_engine: xelatex
    df_print: paged
    fig_caption: yes
  word_document: default
graphics: yes
header-includes:

- \usepackage{fontspec}
- \setmainfont{Arial}

number_sections: yes

geometry: margin = 1.25 cm
documentclass: article
fontsize: 11 pt

fig_width: 5 
fig_height: 3 
fig_caption: true
---
\newpage 

```{r setup , include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', cache=FALSE, cache.lazy = FALSE)
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_knit$set(progress = FALSE, verbose = FALSE)
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
options(knitr.duplicate.label = "allow")
```

All the library were used for analyzing of this dataset as follows:

```{r installed the required packages, echo=TRUE,include=TRUE}
if(!require(pacman))install.packages("pacman")
pacman::p_load(
  tidyverse,
  dplyr,
  ggplot,
  caret,
  magnittr,
  pacman,
  GGally,
  knitr,
  parallel, 
  rattel,
  tictoc,
  gridExtra,
  kableExtra,
  readr, 
  purrr,
  randomForest,
  pROC,
  fastDummies, 
  rpart.plot,
  data.table, 
  reshape2,
  graphics,
  corrplot,
  latexpdf,
  ReporteRs,
  tinytex, 
  latexdiffr,
  latex2exp,
  ggrepel,
)
```

The US murders dataset is included as part of the "dslabs" package.
 
```{r, message=FALSE,echo=TRUE,include=TRUE }
filename <- "murders.csv"
dir <- system.file("extdata", package = "dslabs")
fullpath <- file.path(dir, filename)
#file.copy(fullpath, "murders.csv")
murder <- read_csv(filename)
```

This dataset is consist of 5 column and 51 rows (Table 1). The name of columns are: State, Abbreviation(Abb), Region, Population and total (as shown in Table 2). There is no missing data in dataset (Table 3). The six first row of dataset is shown in Table 4:

```{r specification of murder dataset , echo= FALSE}
kable(dim(murder),
      "pandoc", 
      caption = "The dimention of murder dataset", 
      align = "c", 
      font_size = 5)

kable(names(murder),
      "pandoc", 
      caption = "The name of columns in murder dataset", 
      align = "c", 
      font_size = 5)

kable(any(is.na(murder)),
      "pandoc", 
      caption = "The number of missing data in murder dataset", 
      align = "c", 
      font_size = 5)

kable(head(murder),
      "pandoc", 
      caption = "The first six rows of murder dataset", 
      align = "c", 
      font_size = 5)

```

The first six rows of state names by order of their total murders as shown in Table 5.

```{r State names based on order of total murder, echo=TRUE, include=TRUE}

ind <- order(murder$total)
State_order_murder <- murder$state[ind]

kable(head(State_order_murder),
      "pandoc", 
      caption = "The State names based on order of Their total          murder", 
      align = "c", 
      font_size = 5)
```

It is shown that the maximum total murder is 1257 in California with population around 37253956 and minimum total murder is 2 in Vermont with population around 625741 (Table 6).

```{r order of total murder of state, echo=TRUE, include=TRUE}
# the max and min of total murder 
max_total_murder <- max(murder$total)
min_total_murder <- min(murder$total)

# The state name of min and max of total number 
i_max <- which.max(murder$total)
State_max_total_murdre <- murder$state[i_max]
i_min <- which.min(murder$total)
state_min_total_murdre <- murder$state[i_min]

# the population of max and min of total number state
Population_max_total_murder <- murder$population[i_max]
Population_min_total_murder <- murder$population[i_min]

kable(rbind(max_total_murder,min_total_murder,
      State_max_total_murdre, state_min_total_murdre,
      Population_max_total_murder, Population_min_total_murder),
      "pandoc", 
      caption = "The first six rows of murder dataset with adding       murder rate column", 
      align = "c", 
      font_size = 5)
```

In order to compare the total murder of each state, it is better defined one variable as murder rate. Murder rate is defined as the total murder to 100000 population of each state. The first six rows of murder dataset is shown in Table 7.  


```{r murder rate , echo= TRUE, include=TRUE}
murder <-murder %>% mutate(region = factor(region),
                             murder_rate = total / population *                               10^5)
kable(head(murder),
      "pandoc", 
      caption = "The first six rows of murder dataset with adding murder rate column", 
      align = "c", 
      font_size = 5)

```

Based on murder rate column, the maximum murder rate is 16.42 per 100000 population for District of Columbia.Also the minimum murder rate is 0.32 per 100000 population for Vermont (Table 8).

```{r murder rate of state, echo=TRUE, include=TRUE}
# The state name of min and max of murder rate
max_murder_rate <- max(murder$murder_rate)
min_murder_rate <- min(murder$murder_rate)

i_max_murder_rate <- which.max(murder$murder_rate)
State_max_murdre_rate <- murder$state[i_max_murder_rate]
i_min_murder_rate  <- which.min(murder$murder_rate)
state_min_murdre_rate <- murder$state[i_min_murder_rate ]

kable(rbind(max_murder_rate,min_murder_rate,
      State_max_murdre_rate, state_min_murdre_rate),
      "pandoc", 
      caption = "The first six rows of murder dataset with adding murder rate column", 
      align = "c", 
      font_size = 5)
```

The distribution of murder rate per 100000 population is shown for each state in Figure 1. 

```{r plot murder rate of state, echo=TRUE, include=TRUE, out.width='50%', fig.align='center', fig.cap='Variation of murder rate based on state'}
murder %>% mutate(abb = reorder(abb,murder_rate)) %>%
  ggplot(aes(abb, murder_rate)) +
  geom_col(width = 0.7, color = "pink",fill = "blue") +
  xlab("Abb") +
  ylab("Murder rate per 100000" ) +
  coord_flip() +
  theme(axis.text.x = element_text(size = 10)) +
  theme(axis.text.y = element_text(size = 4))
```

The histogram of murder rate shows that there is a wide range of values with most of them between 2 and 3 and one very extreme
case with a murder rate of more than 15 (Figure 2). 

```{r Distribution of murder rate, echo=FALSE, include=TRUE, out.width='50%', fig.align='center', fig.cap='Distribution of murder rate'}
murder %>% ggplot(aes(murder_rate)) + 
  geom_histogram(bins = 7, color = "pink", fill = "blue") +
  xlab("Murder rate") +
  ylab("Frequency")
```

The murder rate and population are compared in different region as well. Figure 3 shows that in South region the murder rate and total murder are higher than other region. although these relation is not correct for other regions.

```{r boxplot of murder rate and total murder, echo=TRUE, include=TRUE, out.width='50%', fig.align='center', fig.cap='Varioation of murder rate and total murder versus region'}

p1_murder_rate <- murder %>% group_by(region) %>% 
  ggplot(aes(region, murder_rate)) +
  geom_boxplot(col = "pink", fill = "blue")+ 
  xlab("Region") + ylab("Murder rate") 
  
p1_population <- murder %>%  
  group_by(region) %>%
  ggplot(aes(region, population/10^6)) +
  geom_bar(aes(fill = region ),stat="identity")+
  xlab("Region") +
  ylab("Population in million")+ 
  theme(legend.position =  "none") +
  scale_y_log10()

grid.arrange(p1_murder_rate,p1_population) 
```

The total murder and murder rate of each state versus population in million at log scale are shown detailed with state at different region (Figure 5). 

```{r point plot of total murder and murder rate based on population, echo=TRUE, include=TRUE, out.width='50%', fig.align='center', fig.cap='Varioation of Murder rate (log scale) and Total murder (Log scale) versus population in million (log scale)'}

p1 <- murder %>% ggplot(aes(population/10^6, total, label = abb)) +
  geom_point(aes(col=region), size = 3) +
  geom_text_repel() +
  scale_x_log10() +
  scale_y_log10() +
  xlab("Populations in millions (log scale)") +
  ylab("Total murders (log scale)") +
  ggtitle("Total Murder versus population in millions") +
  scale_color_discrete(name = "Region")+
  facet_grid(. ~ region) +
  theme(legend.position = "none")

p2 <- murder %>% ggplot(aes(population/10^6, murder_rate, label = abb)) +
  geom_point(aes(col=region), size = 3) +
  geom_text_repel() +
  scale_x_log10() +
  scale_y_log10() +
  xlab("Populations in millions (log scale)") +
  ylab("Murder rate (log scale)") +
  ggtitle("Murder rate versus Population in millions") +
  scale_color_discrete(name = "Region")+
  facet_grid(. ~ region) +
  theme(legend.position = "none")

grid.arrange(p1,p2)

```


